Parallel for Oregon Zip and WRF_Grid Intersection Proportion

### 2017-03-27 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
The whole U.S shapefile can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
<br/>The Oregon zip code can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

@@

## Steps

@@

<ul class="incremental"> 
  <li>One choice: <br/>  simple calculation; copy and paste</li> 
  <li>Another choice: <br/>  for loop</li> 
</ul>

@@

- The apply() Function
- The lapply() Function
- The sapply() Function
- The tapply() Function
- The mapply() Function

## Parallel on server

```{r cars, echo = TRUE}
summary(cars)
```

>>>>

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
<br/>The whole U.S shapefile can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
<br/>The Oregon zip code can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

@@

## Steps of calculating the proportion of a specific zip code polygon and WRF Grid.
1. Choose zip code = 94705. WRF Grid = 261 and 262. (6 WRF Grid in total. Only 
two are chosen for test)
```{r test, warning= FALSE}

test_zip <- c(97405)
test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]

plot(test_zip_map)
plot(smoke_grid, add = T)
invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
               labels=as.character(smoke_grid$WRFGRID_ID)))

```

```{r include = FALSE}
test_grid <- over(smoke_grid, test_zip_map)
summary(test_grid) # 6 grids over the zip code
test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
test_grid2
```

2. Show the actual location in the map.
```{r}
# test zip map over zip grid
plot(smoke_grid)
plot(test_zip_map, add = T)
zip_over_grid <- over(test_zip_map, smoke_grid)
# summary(zip_over_grid) # this way retains the values
```

```{r include = FALSE}
# test_zip <- c(97405)
# test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]

# plot(test_zip_map)
# plot(smoke_grid, add = T)
# invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
#                labels=as.character(smoke_grid$WRFGRID_ID)))

shape_zip <- SpatialPolygons(test_zip_map@polygons)
shape_grid <- SpatialPolygons(smoke_grid@polygons)
plot(shape_zip)
plot(shape_grid, add = T)
# try and plot the values for this zipcode
invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
               labels=as.character(smoke_grid$WRFGRID_ID))) # grid width wider than previous map
```

```{r include = FALSE}
# output the WRF Grid 261 and 262 for test
wrf_grid_261 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 261, ]
plot(wrf_grid_261)

wrf_grid_262 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 262, ]
plot(wrf_grid_262)
```

```{r include = FALSE}
# calculate the area of grid 261 and 262
gArea(SpatialPolygons(wrf_grid_261@polygons))
gArea(SpatialPolygons(wrf_grid_262@polygons))
```

## Steps (Continued)
3. Show the intersection of chosen WRF Grid and zip code.
```{r show, warning=FALSE}
# Subset to one zip code and two different WRF grids ---------------------------
# plot zip and 2 grids
plot(test_zip_map)
#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
#               labels=as.character(test_zip_map$ZCTA5CE10)))
plot(wrf_grid_261, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))
plot(wrf_grid_262, add=T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
               labels=as.character(wrf_grid_262$WRFGRID_ID)))
```

4. The proportion of two WRF Grid.
```{r calculation, warning=FALSE}
# Start with the intersection with wrf grid 261 and zipcode
# first I need to convert the spatial polygon to just polygon
poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

zip_261_int <- gIntersection(poly_261, shape_zip)
plot(zip_261_int)
plot(wrf_grid_261, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))

prop_261_int <- gArea(zip_261_int)/gArea(poly_261)
prop_261_int # 41.8% of grid is covered by zip

# now what about grid 262
poly_262 <- SpatialPolygons(wrf_grid_262@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

zip_262_int <- gIntersection(poly_262, shape_zip)

plot(zip_262_int)
plot(wrf_grid_262, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))

prop_int_262 <- gArea(zip_262_int)/gArea(poly_262)
prop_int_262 # 6.0% of grid is covered by zip
```

@@

<ul class="incremental"> 
  <li> One choice: <br/>  simple calculation; copy and paste</li> 
  <li> Another choice: for loop
  <br/> When analyzing big data set (like millions of observations), it is 
  unrealistic to copy so many rows of similar or same code.</li><br/>
  <li> And... <br/> Why using apply function? </li>
  <li> Code Readability <br/>  Understand the purporse more directly.</li> 
  <li> Working Efficiency <br/>  When analyzing big data set (like millions of 
  observations), it is unrealistic to copy so many rows of similar or same code.</li> 
</ul>

@@

- The apply() Function
<br/> Apply functions over array margins.<br/>

- The lapply() Function
<br/> Apply a function over a list or vector.<br/> 
Apply a given function to every element of a list and obtain a list as result.<br/>

- The sapply() Function
<br/> Apply a function over a list or vector.<br/>
Works like lapply(), but it tries to simplify the output to the most elementary data structure that is possible.

- The tapply() Function
<br/> Apply a function over a ragged array.

- The mapply() Function
<br/> Apply a function to multiple list or vector arguments.

## Parallel on server
One core vs Multiple cores

```{r}

for(i in 1:length(or_zip_name)){
  # output value of zipcode
  zipcode <- as.character(or_zip_name[i]) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)

  # now I can create the second loop that finds the proportion of the area of
  # the zipcode polygon that overlaps with each WRF-Grid
  for(j in 1:length(wrf_grid_name)){
    # output each grid and create a polygon
    wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == j, ]
    # now what about grid 719; should be much less
    wrf_poly <- SpatialPolygons(wrf_grid@polygons)

    zip_wrf_intersect <- gIntersection(wrf_poly, zip_poly)
    # if empty, then set to 0, else find the proportion
    grid_prop <- ifelse(is.null(zip_wrf_intersect),
                        0, gArea(zip_wrf_intersect)/gArea(wrf_poly))
    # populate the matrix based on i position and j position
    zip_wrf_proportion[[i,j]] <- grid_prop
  }
}

```

```{r, eval=FALSE}
f1 <- function(x){
  zipcode <- as.character(x) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)
}
zip_polygon <- lapply(or_zip_name, f1) # list
# b <- zip_polygon[1] # list
# c <- zip_polygon[[1]] # spatial polygon

vars1 <- c(1:1610)
f2 <- function(n){
  wrf_grid <- smoke_grid[smoke_grid@data$WRFGRID_ID == n,] # sp df
  # now what about grid 719; should be much less
  wrf_poly <- SpatialPolygons(wrf_grid@polygons) # sp
}
wrf_polygon <- lapply(vars1, f2) # list of 1610 wrf_grid

f3 <- function(x,y){
  zip_wrf_intersect <- gIntersection(wrf_polygon[y][[1]], zip_polygon[x][[1]])
  grid_prop <- ifelse(is.null(zip_wrf_intersect),
                      0, gArea(zip_wrf_intersect)/gArea(wrf_polygon[y][[1]]))

}
proportion <- mcmapply(f3, rep(1:417, each=1610), rep(1:1610,417))

```
